\large Neural, Tree, and Symbolic Optimization of Cosmological Particle-Mesh Simulation \vspace{0.3em}

### 2023-06-26 docs/papers/sto/so.tex
FIXUP draft a few sentences on so training

Dimensionless combinations, e.g.,
%
\begin{equation}
f(k_i; \vvartheta) = f(k_i/k_P, kR_\tophat, \cdots; D, f, \cdots).
\end{equation}
%
$f$ and $g$ are optimized during training to correct and sharpen the
gravitational forces.
Gravity obeys the $\mathrm{O}(3)$ rotational symmetry, which is however
broken by the PM solver to the O\textsubscript{h} discrete symmetry with
48 elements involving permutations and reversals of the 3 principal axes
of the simulation box.
The PM evaluation of each force component can further break the symmetry
along that axis, leaving the remaining 2 permutationally symmetric.
Therefore, we implement $f$ and $g$ to be even functions of the
wavevector components, and $g$ to be permutation equivariant in all 3
components, so that $f g$ is the most general modification to the force
kernel.
An even and real kernel in configuration space is also even in Fourier
space.
We take the absolute value for the evenness, and sort the 3 wavevector
compontents to achieve the permutation equivariance.

\YZ{
The parameters of the neural networks, which enter every force evaluation in
every time integration step, can be treated just the same as other cosmological
parameters.
The objective gradients are accumulated, based on which the parameters are then
optimized.
}

@@

We perform parallel training using distributed GPU devices.
The training on each snapshot is performed on a single GPU device.
The gradients are collected and averaged to update the SO parameters.

>>>>

Dimensionless combinations, e.g.,
%
\begin{equation}
f(k_i; \vvartheta) = f(k_iR_d, k_iR_\tophat, \cdots; G_m, \Omegam, \cdots)
\end{equation}
%
\begin{equation}
g(k_1, k_2, k_3; \vvartheta) = g(\{k_jR_d, k_jR_\tophat, \cdots\}_{j=1,2,3}; G_m, \Omegam, \cdots)
\end{equation}
%
$f$ and $g$ are optimized during training to correct and sharpen the
gravitational forces.
Gravity obeys the $\mathrm{O}(3)$ rotational symmetry, which is however
broken by the PM solver to the O\textsubscript{h} discrete symmetry with
48 elements involving permutations and reversals of the 3 principal axes
of the simulation box.
The PM evaluation of each force component can further break the symmetry
along that axis, leaving the remaining 2 permutationally symmetric.
Therefore, we implement $f$ and $g$ to be even functions of the
wavevector components, and $g$ to be permutation equivariant in all 3
components, so that $f g$ is the most general modification to the force
kernel.
An even and real kernel in configuration space is also even in Fourier
space.
We take the absolute value for the evenness, and sort the 3 wavevector
compontents to achieve the permutation equivariance.

\YZ{
The parameters of the neural networks, which enter every force evaluation in
every time integration step, can be treated just the same as other cosmological
parameters.
The objective gradients are accumulated, based on which the parameters are then
optimized.
For both $f$ and $g$ neural networks, we adopt the multilayer perceptron (MLP)
structure, each with one input layer, two hidden layers and one output layer.
}

@@

We perform parallel training using distributed GPU devices.
The training on each snapshot is performed on a single GPU device.
During each training step, the gradients are collected and averaged across
devices to update the SO parameters.
